### 2020-08-27 README.Rmd
Moved list of journals from paper.Rmd to README. changed "Chinook salmon" to "Chinook Salmon" to match AFS guidelines.

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
# Please put your title here to include it in the file below.
Title <- "State-Space Model to Estimate Salmon Escapement Using Multiple Data Sources"
```

**Possible journals:**

First Choice:

* **[North American Journal of Fisheries Management](https://afspubs.onlinelibrary.wiley.com/journal/15488675)**

Second Choice:

* [Transactions of the American Fisheries Society](https://afspubs.onlinelibrary.wiley.com/journal/15488659)

Others:

* [Canadian Journal of Fisheries and Aquatic Sciences](http://www.nrcresearchpress.com/page/cjfas/editors)

* [Fisheries Research](https://www.journals.elsevier.com/fisheries-research)

* [Fishery Bulletin](http://fishbull.noaa.gov/)

* [Fish and Fisheries](http://onlinelibrary.wiley.com/journal/10.1111/(ISSN)1467-2979/homepage/ProductInformation.html)

# ManuscriptSTADEM

[![Binder](https://mybinder.org/badge_logo.svg)](https://mybinder.org/v2/gh/KevinSee/ManuscriptSTADEM/master?urlpath=rstudio)

This repository contains the data and code for our paper:

> See, K.E., R.N. Kinzer and M.W. Ackerman (2020). _`r Title`_. Name of journal/book <https://doi.org/xxx/xxx>

Our pre-print is online here:

> See, K.E., R.N. Kinzer and M.W. Ackerman (2020). _`r Title`_. Name of journal/book, Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/xxx/xxx>

### How to cite

Please cite this compendium as:

> See, K.E., R.N. Kinzer and M.W. Ackerman (`r format(Sys.Date(), "%Y")`). _Compendium of R code and data for `r Title`_. Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/xxx/xxx>

### How to download or install

You can download the compendium as a zip from from this URL: <https://github.com/KevinSee/ManuscriptSTADEM/archive/master.zip>

Or you can install this compendium as an R package, ManuscriptSTADEM, from GitHub with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
remotes::install_github("KevinSee/ManuscriptSTADEM")
```

To

### Licenses

**Text and figures :**  [CC-BY-4.0](http://creativecommons.org/licenses/by/4.0/)

**Code :** See the [DESCRIPTION](DESCRIPTION) file

**Data :** [CC-0](http://creativecommons.org/publicdomain/zero/1.0/) attribution requested in reuse

### Contributions

We welcome contributions from everyone. Before you get started, please see our [contributor guidelines](CONTRIBUTING.md). Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). By participating in this project you agree to abide by its terms.