### 2021-03-09 analysis/paper/paper.Rmd
started to address reviewer comments

---
title: "State-Space Model to Estimate Salmon Escapement Using Multiple Data Sources"
author:
  - Kevin E. See:
      email: Kevin.See@merck.com
      institute: [BM]
      correspondence: true
  - Ryan N. Kinzer:
      email: ryank@nezperce.org
      institute: [NezPerce]
      correspondence: false
  - Michael W. Ackerman:
      email: mike.ackerman@merck.com
      institute: [BM]
      correspondence: false
institute:
  - BM: Biomark, Inc., 705 South 8th Street, Boise, Idaho, 83702, USA
  - NezPerce: Nez Perce Tribe, Department of Fisheries Resource Management, 14054 Burr Road, PO Box 1942, McCall, Idaho, 83638, USA
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
    bookdown::word_document2:
      fig_caption: yes
      reference_docx: "../templates/template.docx" # Insert path for the DOCX file
      number_sections: FALSE
      toc: no
      always_allow_html: true
      fig_width: 6
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks.lua
      - --lua-filter=../templates/pagebreak.lua
    bookdown::html_document2:
      fig_caption: yes
      number_sections: FALSE
      toc: no
      fig_width: 7
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks2.lua
      - --lua-filter=../templates/pagebreak.lua
    bookdown::pdf_document2:
      fig_caption: yes
      number_sections: FALSE
      toc: no
      geometry: margin=1in
      latex_engine: "xelatex"
      fig_width: 7
      # fig_height: 6
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks2.lua
      - --lua-filter=../templates/pagebreak.lua
      includes:
        in_header: ../templates/header_manuscript.tex
fontsize: 12pt
mainfont: Times New Roman
bibliography:
- references.bib
- packages.bib
csl: "../templates/american-fisheries-society.csl" # Insert path for the bib-style
abstract: |
  Accurate estimates of salmonids passing Lower Granite Dam on the Snake River, by species and origin, are a critical input to assessing the status and trends of various populations, as well as successful management of fisheries in the Snake River basin. Here we describe a state-space model that estimates such escapement past a dam by using window counts, passive integrated transponder (PIT) tag observations and data from an adult fish trap, accounting for issues such as nighttime passage, fallback and re-ascension, potential observation error at the window and uncertainty in the adult trap rate. We tested the approach using a simulation framework that mimicked several levels of observation error, differences between nighttime passage and re-ascension rates and the possibility of the adult trap being closed for some period of time. Our results demonstrate that the model produced unbiased estimates across all tested scenarios. We also applied this model to data from the Lower Granite Dam to produce estimates of wild, clipped hatchery and unclipped hatchery spring/summer-run Chinook Salmon and steelhead from spawn years 2010-2019. 
keywords: |
  escapement; Chinook; steelhead; salmon; Lower Granite Dam; state-space model; Bayesian model
highlights: |
  These are the highlights. 
---

```{r setup, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  comment = "#>",
  fig.path = "../figures/",
  dpi = 300
)

# set default values for saving alternative figures (color versions)
my_dpi = knitr::opts_chunk$get()$dpi
my_ext = "tiff"
```

```{r load_libraries}
# setwd('analysis/paper')

library(tidyverse)
library(lubridate)
# library(msm)
# library(HDInterval)
library(magrittr)
library(ggpubr)
library(knitr)
library(kableExtra)

library(STADEM)

# set default theme for ggplot2
theme_set(theme_pubr())

# default options for tables
options(knitr.kable.NA = '-')

# when knitting to Word, use this
# what kind of document is being created?
doc.type <- knitr::opts_knit$get('rmarkdown.pandoc.to')

if(doc.type == 'docx') {
  options(knitr.table.format = "pandoc")
}

if(doc.type %in% c("docx", "html")) {
  theme_set(theme_pubr(base_family = "Times New Roman"))
}

```

```{r package-bibtex, eval = F}
knitr::write_bib(c("lubridate",
                   "msm",
                   "HDInterval",
                   "ggpubr",
                   "knitr", 
                   "rjags",
                   "rmarkdown"),
                 file = 'packages.bib')
```

```{r load_sim_results, cache = F}

result_nms <- list.files('../data/derived_data/SimulationFits')
result_nms <- result_nms[!grepl('other', result_nms)]
scenario_nms <- gsub('^Sim_', '', result_nms)
scenario_nms <- gsub('\\.rda$', '', scenario_nms)
# length(scenario_nms)

# compile all results into single dataframe
res_df <- scenario_nms %>%
  as.list() %>%
  rlang::set_names() %>%
  map_df(.id = 'Scenario',
         .f = function(nm) {
    load(paste0('../data/derived_data/SimulationFits/Sim_', nm, '.rda'))
    res %>%
      map_df(.id = 'sim',
             .f = identity)
  }) %>%
  mutate(Variable = factor(Variable, levels = c('Unique.Wild.Fish',
                                                'Unique.Hatch.Fish',
                                                'Unique.HNC.Fish',
                                                'All.Fish',
                                                'Daytime.Fish',
                                                'Night.Fish',
                                                'Reascent.Fish'))) %>%
  mutate(Scenario = recode(Scenario,
                           'Baseline_WinErr_high' = 'Baseline_Err_H',
                           'Baseline_WinErr_low' = 'Baseline_Err_L',
                           'NightReasc' = 'N-R',
                           'NightReasc_WinErr_high' = 'N-R_Err_H',
                           'NightReasc_WinErr_low' = 'N-R_Err_L',
                           'TrapShutdown' = 'TrapDown',
                           'TrapShutdown_WinErr_high' = 'TrapDown_Err_H',
                           'TrapShutdown_WinErr_low' = 'TrapDown_Err_L',
                           'NightReasc_TrapShutdown' = 'N-R_TrapDown',
                           'NightReasc_TrapShutdown_WinErr_high' = 'N-R_TrapDown_Err_H',
                           'NightReasc_TrapShutdown_WinErr_low' = 'N-R_TrapDown_Err_L'),
         Scenario = factor(Scenario, 
                           levels = paste0(rep(c('Baseline', 'TrapDown', 'N-R', 'N-R_TrapDown'), each = 3), 
                                           c('', '_Err_L', '_Err_H'))))

#------------------------------------------------------
# Re-format res_df into long format of point ests.,
# CI's and CIin for both Models : STADEM, SCOBI
#------------------------------------------------------
res_long_df <- res_df %>%
  pivot_longer(cols = -(Scenario:Truth),
               names_to = "key",
               values_to = "point") %>%
  separate(key, into = c("Model","est"), sep = "_") %>%
  pivot_wider(names_from = "est",
              values_from = "point") %>%
  mutate(Model = recode(Model, 
                        "ISEMP" = "STADEM")) %>%
  rowwise() %>%
  mutate(se = (uppCI - lowCI) / (2 * qnorm(0.975)),
         cv = se / est,
         bias = est - Truth,
         rel_bias = bias / Truth,
         MSE = mean((bias)^2),
         inCI = ifelse(inCI == 1, T, ifelse(inCI == 0, F, NA)),
         Window = ifelse(grepl("H",Scenario),"Window_Err_High",
                         ifelse(grepl("L",Scenario),"Window_Err_Low","No Window Error")),
         Night = ifelse(grepl("N-R",Scenario),"Yes","No"))
```

```{r load-lgr-results}
lgr_res = list.files('../data/derived_data/STADEM_results') %>%
  as.list() %>%
  map(.f = function(x) {
    load(paste('../data/derived_data/STADEM_results', x, sep = '/'))

    tot_est = stadem_mod$summary %>%
      as_tibble(rownames = 'variable') %>%
      filter(grepl("X.tot.new", variable)) %>%
      mutate(cv = sd / mean) %>%
      mutate(Year = str_extract(x, "[:digit:]+") %>%
               as.numeric()) %>%
      select(Year, variable, mean, cv) %>%
      bind_rows(stadem_list$weeklyData %>%
                  group_by(Species) %>%
                  summarise_at(vars(mean = win_cnt),
                               list(sum),
                               na.rm = T) %>%
                  mutate(variable = "win_cnt")) %>%
      tidyr::fill(Year, Species,
                  .direction = "downup") %>%
      select(Species, Year, everything())

    rate_df = stadem_list$weeklyData %>%
      mutate(Year = str_extract(x, "[:digit:]+") %>%
               as.numeric()) %>%
      select(Species, Year, Start_Date:reascent_tags) %>%
      mutate(night_rate = 1 - day_tags / tot_tags,
             reasc_rate = reascent_tags / tot_tags)

    return(list(tot_est = tot_est,
                rate_df = rate_df))

  })

lgr_est = map_df(lgr_res,
                 .f = "tot_est")

rate_comp = map_df(lgr_res,
                   .f = "rate_df")

```

# Introduction

Fish escapement refers to the number of adults that survive juvenile and subadult rearing, escape harvest, and achieve a size and age to potentially spawn [e.g., @Bue1998]. For anadromous fishes, escapement is often estimated at a fixed location in a river system prior to fish reaching their natal habitat to spawn. Escapement estimates facilitate effective fisheries management, particularly estimates of escapement for component groups [e.g., by stock, population, age, origin [wild, hatchery], @Hess2014; @Steinhorst2017; @Camacho2018] which provide valuable information that fisheries managers use to achieve sustainable harvest, while protecting small and vulnerable populations. Accurate escapement estimates are increasingly important for depleted populations as they inform status metrics [@McElhany2000] and facilitate assessments of population viability, extinction risk [@NWFSC2015; @Williams2016] and climate change vulnerability [@Crozier2019].

Populations of Chinook Salmon *Oncorhynchus tshawytscha* and steelhead trout *O. mykiss* in the Snake River basin of Pacific Northwest, USA, are depleted following decades of substantial harvest and anthropogenic changes to their migration corridor (e.g., construction of hydroelectric projects on the Snake and Columbia rivers) and tributary habitats [@Nehlsen1991; @McClure2003; @NWFSC2015]. As a result, Snake River spring/summer-run Chinook Salmon (hereafter sp/sum Chinook Salmon) were classified as threatened in 1992 under the Endangered Species Act (ESA; Federal Registry Notice 57 FR 14653), and Snake River summer-run steelhead trout (hereafter steelhead) were listed as threatened five years later (Federal Registry Notice 62 FR 43937). Snake River sp/sum Chinook Salmon and steelhead have substantial cultural, recreational, commercial and subsistence value within the Snake River basin as well as in downstream corridors (i.e., Columbia River) and ocean fisheries. The aggregate escapement of Snake River sp/sum Chinook Salmon and steelhead populations, with the exception of Tucannon River (a tributary of the Snake River whose confluence is downstream of Lower Granite Dam), is monitored at Lower Granite Dam located in southeast Washington; the final dam on the Snake River that returning adults must pass prior to heading to tributary spawning locations. Many fisheries management and conservation actions are made based on estimates of escapement at Lower Granite Dam parsed by species and origin [@NMFS2019; @NWFSC2015; @NPCC2014]. Additionally, harvest openings and closures, both upstream in Snake River fisheries and downstream in mainstem Snake and Columbia rivers fisheries, are predicated on escapements at Lower Granite Dam.

The majority of sp/sum Chinook Salmon and steelhead returning to the Snake River must ascend a fish ladder on Lower Granite Dam before migrating upriver. Fish management agencies previously used counts of fish, by species, passing an observation window within the fish ladder as a census of fish escaping past the dam. For both species, total escapement was then parsed into groups (e.g., wild, hatchery) using observed marks and genetic data from a sample of fish captured at an adult trap located on the fish ladder upstream of the observation window [@Camacho2018; @Steinhorst2017; @Steele2019]. Treating window counts as a census proved beneficial as being an easy, straight-forward method ascertained in near real-time. Moreover, downriver fisheries management arenas have used window counts at lower-river dams as escapement estimates for the past several years, and consistency in methods is often desirable for management decision making [@WDFW2019].

However, using window counts as a census of sp/sum Chinook Salmon and steelhead passing Lower Granite Dam can be problematic as it fails to account for multiple sources of uncertainty and disregards known biological processes. In-person window counts, where observers look directly into the fish ladders to identify and count all passing fish, by species, occur for 50 minutes per hour, 16 hours a day from April through October (which corresponds with peak run timing for several species). Counts are then expanded to provide an estimate for the entire hour [@USACE2015]. For the remainder of the year (November through March), fish passage is video-taped for 10 hours each day; fish counters then read the video tapes and submit daily fish counts. Typically, the observational error rates of live and video window counts are unknown, and sampling error rates are ignored. @Hatch1994 conducted a study of the counting system at Lower Granite Dam in 1992, and found that sp/sum Chinook Salmon were undercounted at the window by in-person window counters compared to video counts. There was a significant difference between daily standard counts (16 or 10 hours per day) and daily 24-hour video counts; they found a significant nighttime passage rate (crossing while the counting window was closed) with 3.5% of sp/sum Chinook Salmon and 6.6% of steelhead adults passing during "nighttime" hours. They also found some species misidentification issues, which could result in either under- or over-counting of a particular species. While that study identified potential issues with the current window counting procedure, it was only conducted in a single year, nearly 30 years ago, at a time when sp/sum Chinook Salmon and steelhead total returns were lower than in the past decade.

One additional biological process, the fallback and re-ascension of adult migrants at the dam, should also be accounted for to estimate the true escapement at Lower Granite Dam. Adult fish that migrate through the ladder and past the dam may fallback over the dam (e.g., over a spillway, through navigation locks) and later may re-ascend the fish ladder again [@Boogs2004] and be double-counted at the window (re-ascension). Whereas unaccounted nighttime passage can lead to an underestimate in escapement, both fallback without and with re-ascension instead potentially lead to an overestimate of escapement [@Dauble2000]. Previously, it was often assumed that these two biological processes canceled each other out, resulting in window counts providing an unbiased estimate of escapement [@Camacho2018].

In this study, we describe a novel method to estimate aggregate and group escapements which incorporate all sources of known uncertainty and demonstrate that the estimation method is essentially unbiased, thus better informing conservation and management decision making. We also test whether observed nighttime passage and fallback/re-ascension rates are typically equal. Our method for estimating species-specific escapement past Lower Granite Dam incorporates window counts, data from the adult fish trap, and observations of fish previously tagged with passive integrated transponder (PIT) tags in the adult ladder to explicitly model nighttime passage, re-ascension, and observation error using a state-space approach [@Royle2008] which separates process variance (e.g., week to week variance in true escapment) from observation error variance (e.g., observation error at the window, or sampling variance at the trap). To meet desired management and conservation objectives, modeled escapement includes estimates of uncertainty and is parsed into weekly strata. Further, total and weekly estimates are parsed into three origin groups: wild fish, hatchery fish with a clipped adipose fin, and unclipped hatchery fish. Estimates of escapement account for fish that migrate through the ladder outside of observation hours (nighttime passage) and those that ascend the ladder multiple times (re-ascension). Our model is implemented in the **ST**ate-space **A**dult **D**am **E**scapement **M**odel (STADEM) package for the statistical software R [@R-Core-Team2020], and is available for download or installation from [https://github.com/KevinSee/STADEM](https://github.com/KevinSee/STADEM). To validate the STADEM results, we simulated 12 scenarios with varying trapping rates, fallback and re-ascension rates, nighttime passage rates, and window count error rates. We then applied this model to sp/sum Chinook Salmon and steelhead returns at Lower Granite Dam for spawn years 2010 - 2019. The STADEM model combines multiple imperfect sources of data to reduce bias in escapement estimates and provides improved estimates of uncertainty.

# Methods

## Data Requirements

We used STADEM and three sources of data to estimate sp/sum Chinook Salmon and steelhead escapement at Lower Granite Dam from 2010-2019. Data sources included 1) counts of fish migrating past the observation window located on the adult fish ladder at Lower Granite, 2) information from adults captured at a fish trap located in the fish ladder, and 3) observations of fish PIT tagged prior to arriving at Lower Granite Dam and detected in the adult fish ladder. Below, we describe each of the data sources in more detail as they pertain to Lower Granite Dam; similar data could likely be obtained from other fish passage facilities.

### Window Counts

Daily counts of adult sp/sum Chinook Salmon and steelhead passing the observation window located on the Lower Granite Dam fish ladder were estimated and provided by the US Army Corps of Engineers. When summed, they provide an estimate of the number of fish ascending and passing Lower Granite Dam each season. Window counts were made for each species using direct in-person visual monitoring and video monitoring during daytime hours [@Hatch1994]. Direct visual monitoring occurred during peak run times (April 1 – October 31) for 16 hours per day (0400 – 2000 hours). Video monitoring occurred during the beginning and tail ends of the adult runs (March 1 – March 31 and November 1 – December) for 10 hours per day (0600 – 1600 hours) [@USACE2015]. During direct visual monitoring, observers recorded each adult (≥ 30cm), by species, passing the window for 50 minutes of each hour of operation. The sum of the daily 50-minute counts were then multiplied by 1.2 to account for the 10 minutes when fish were not counted. Daytime window counts were not expanded for fish that may have ascended the ladder outside of operational hours (i.e., nighttime passage) [@USACE2015]. Window counts were accessed through the Columbia Basin Research Data Access in Real Time (DART) website, [www.cbr.washington.edu/dart/query/adult_daily](www.cbr.washington.edu/dart/query/adult_daily), using their window count query. Counts were provided for each day the fish ladder was open to passage.

### Adult Fish Trap Data

The second source of data are from a sample of fish collected in the adult trap as they migrated past Lower Granite Dam [@Ogden2016b]. The trap, also located within the adult fish ladder and upstream of the observation window, provides an opportunity to collect biological data (e.g., origin [wild, hatchery], genetic stock, length, age, sex) from captured adults which allows for decomposition of the escapement into specific groups [e.g., @Camacho2018; @Steinhorst2017]. The trap was operational for 24 hours per day and randomly sampled the run by opening four times per hour for a length of time determined by a set daily trapping rate. The trap rate was determined by a committee of collaborating management agencies with a goal of capturing a target number of wild fish while also balancing fish handling concerns. Trap sample rates were typically 10-25%, but fluctuated throughout the season due to, for example, high water temperatures, decreased flows, trap malfunctions and/or closures, fish handling logistics, in-season forecast adjustments, or adjustments due to shifting species composition throughout the year.

All captured fish were anesthetized, identified to species, examined for existing marks or tags, measured for fork length and visually identified as wild or hatchery. The most widely used marking of hatchery fish is a clipped (removed) adipose fin, although coded wire tags are used in less than 10% of the hatchery releases. Some subset of hatchery fish are either intentionally or unintentionally released without a clipped adipose fin, and these are referred to as unclipped hatchery fish, or hatchery no-clips (HNC). For adults with intact adipose fins (unclipped), which includes wild and hatchery no-clip individuals, either a portion or all of fish trapped (depending on the year) had scale and genetic tissue samples taken. Scale samples were used to estimate age [@Wright2015] and genetic tissue samples were used to determine sex [@Campbell2012] and estimate the location of origin of wild fish using genetic stock identification [e.g., @Hargrove2019]. Prior to 2013, only fish determined to be wild in origin at the trap were sampled for scale and genetics. Starting in 2013, every unclipped sp/sum Chinook Salmon and steelhead trapped at Lower Granite Dam was genotyped to simplify collaborative logistics and better estimate the proportion of unclipped hatchery fish that appear phenotypically wild. @Camacho2018 provide further details on trap sample rates and valid sample selection. Prior to release, all non-PIT tagged fish with an intact adipose fin (i.e., putatively wild) received a PIT tag. Final determination of wild, clipped hatchery, or unclipped hatchery origins were assigned using a post-hoc analysis of marks and tags, including parentage-based tagging [@Steele2013; @Steele2019]. Data from the adult trap were collected and managed by multiple agencies and were made available by the Idaho Department of Fish and Game [@Camacho2018].

### PIT Tag Data

The last source of data were observations of PIT tagged adult sp/sum Chinook Salmon and steelhead at detection sites located in the Lower Granite Dam fish ladder. These observations provide daily estimates of 1) a trapping rate, 2) a nighttime passage rate, and 3) a re-ascension rate. Detections used in the model included all fish that were previously PIT tagged as juveniles or adults prior to arriving at Lower Granite Dam (i.e., does not include newly tagged adults at the dam) and detected at adult detection sites in the dam passage system. PIT tag data were provided through DART and the adult ladder PIT tag query; [http://www.cbr.washington.edu/dart/query/pitadult_obsyr_detail](http://www.cbr.washington.edu/dart/query/pitadult_obsyr_detail).

A trap rate estimate was derived using Lincoln-Peterson mark-recapture methods [@Seber2002] and PIT tag observations of both sp/sum Chinook Salmon and steelhead at Lower Granite Dam adult detection sites. The "mark" group included all tags (both species) detected in the adult trap and the "capture" group included tags observed to cross the weir at the upstream end of the fish ladder as adults left the passage system. The proportion of tags detected leaving the passage system that were also caught in the trap each week was assumed to reflect the trap rate experienced by target species as they migrated through the ladder. The sample size of previously tagged fish detected at Lower Granite Dam influences the uncertainty in that trap rate, which is also informed by the number of adults caught in the trap and the window counts. The set trapping rate (i.e., the recorded time that the trap is open to trap adults) does not always reflect the true trapping rate or proportion of fish that are actually captured in the trap due to various issues including trap malfunctions, separation-by-code fish opening the trap more frequently than expected, and process error, among others. Therefore, we use the mark-recapture approach to estimate a "true" trapping rate.

We used PIT tag observation data from the adult fish ladder to estimate both nighttime passage and re-ascension rates; each estimated on a weekly basis. The nighttime passage rate was based on the count of PIT tags that migrated through the fish ladder during non-window observation hours and the total number of tags passing the fish ladder. The re-ascension rate incorporated the count of tags observed passing the upstream most detection sites in the adult fish ladder (i.e., passing the dam) and later detected re-entering the downstream end of the fish ladder at a later time and the total number of tags leaving the fish ladder. Previously, we looked for differences in nighttime passage and re-ascension rates estimated using wild fish only, versus combining hatchery and wild fish together, and found no difference. Therefore, we combine wild and hatchery PIT tagged fish observations to estimate common nighttime passage and re-ascension rates to increase sample sizes.

## Model Framework

We estimated the total number of fish crossing the dam each week, based on the window counts and the total fish passing the adult trap, while also accounting for estimated nighttime passage and fallback/re-ascension rates using a state-space modeling approach [@Royle2008]. Our model is implemented in the STADEM package for the R statistical software [@R-Core-Team2020]. We assumed that the window counts and the estimates from the trap (fish in the trap divided by trap rate that week) were generated by processes with observation error. In the case of the trap, for example, we assumed there was sampling variation and uncertainty around our estimates of the true unknown trap rate. STADEM adjusted the window counts for the nighttime passage rate, and both window and trap estimates for the re-ascension rate to estimate the number of unique fish that crossed the dam. Finally, adult sampling data from the trap (wild, hatchery, hatchery no-clip) were used to partition the total escapement estimate by origin (Figure \@ref(fig:stadem-examp-fig)). Additional model details can be found in [Appendix A](#append1). The STADEM package is available from the primary author at [https://github.com/KevinSee/STADEM](https://github.com/KevinSee/STADEM), and requires the use of the JAGS software [@R-rjags] for Bayesian inference.

## Simulations

We tested the STADEM model on a variety of simulated data sets. These simulated data sets contained a fixed number of unique adult fish of known origin crossing a dam, from a total of 25 fictional populations with differential run-timing (i.e., date of passage at Lower Granite Dam). Each simulated fish was given a date of ladder ascension, based on its population and the range of observed run timing for that population. Each fish was also simulated to cross the dam either while the window was open for counting, or not, and was given the chance to be "caught" in the simulated fish trap given the week when it ascended the dam, and the known trap rate that week. Fallback and re-ascension behavior was also simulated, with each fish having the possibility of falling back and re-ascending the ladder up to three times.

Our objective was to examine STADEM model estimates of origin-specific (wild, hatchery, hatchery no-clip) escapement from the combinations of two trap rate scenarios (constant and shut down for 3 weeks), two fallback/re-ascension and nighttime passage combinations, and three window count error rates; resulting in twelve different scenarios (Table \@ref(tab:lgr-sim-tab)). The simulation parameters such as proportion of origin, run-timing, nighttime passage rates, fallback and re-ascension rates and trap rates were based on observed values at Lower Granite Dam between 2010-2015. We generated 99 simulations for each scenario, and ran STADEM on each one, resulting in a total of 1,188 model runs. Further details about simulation procedures can be found in [Appendix B](#append2).

We grouped the simulation results by fish origin and evaluated them by several measures. Relative bias is the difference between the simulated "truth" and the STADEM estimate, divided by the simulated "truth". We estimated the precision by examining the average coefficient of variation (CV) of the estimates. We calculated the root mean squared error (RMSE) as the square root of the mean of the squared bias in the estimate. Finally, we evaluated coverage probabilities by determining what proportion of the model results generated an estimated 95% credible interval that contained the simulated true value.

## Lower Granite Application

We applied STADEM to empirical data from Lower Granite Dam for sp/sum Chinook Salmon and steelhead returning to the Snake River during spawn years 2010 to 2019. Window counts for both species were accessed from DART via functions within STADEM. For sp/sum Chinook Salmon, a spawn year refers to adults that migrate past the dam prior to August 17 each year and spawn that late summer and fall. For steelhead, the spawn year is defined as steelhead that migrate past Lower Granite Dam starting July 1 the previous year and prior to June 30 of the given year (e.g., spawn year 2017 steelhead migrate past Lower Granite between July 1, 2016 and June 30, 2017). Data from the adult trap were made available by Idaho Department of Fish and Game, and adult PIT tag detection data within the fish passage ladder at Lower Granite Dam were accessed from the PTAGIS regional database ([https://www.ptagis.org/](https://www.ptagis.org/)).

All data used in this manuscript, as well as associated code for both the simulations and the Lower Granite appllication, can be found at [https://www.github.com/KevinSee/ManuscriptSTADEM]([https://www.github.com/KevinSee/ManuscriptSTADEM]).

# Results

## Simulations

The STADEM performance evaluation statistics were very similar for hatchery clipped (N = 70,000), hatchery no-clip (N = 5,000) and wild fish (N = 25,000). In the interest of brevity, we only present the results from wild fish, which corresponds to a medium sized escapement level.

STADEM results were very similar across all scenarios (Figure \@ref(fig:rel-bias-lgd)). Estimates of wild escapement were unbiased, with an average relative bias of 0.2–0.3%. The CV of the estimates averaged 2.0–3.0%, with higher CV's being associated with scenarios when the trap was closed for 3 weeks. The coverage probabilities always exceeded 95% across all scenarios. The RMSE was near 500 for each scenario, representing an estimate within 2% of the true value, demonstrating the accuracy of STADEM (Table \@ref(tab:summ-tab)).

## Lower Granite Application

We applied STADEM to data from Lower Granite Dam for sp/sum Chinook Salmon and steelhead for spawn years 2010–2019. Estimates of total escapement, as well as estimates of wild, hatchery clipped, and hatchery no-clip estimates are presented in Table \@ref(tab:lgr-tab). Estimates of total unique fish escaping past Lower Granite Dam were sometimes higher and sometimes lower than the raw window counts, indicating that the relative strength of nighttime passage and re-ascension rates differed across years. CVs ranged from 2.5-7.1% for wild fish, 2.3-5.1% for clipped hatchery fish, 3.4-8.8% for hatchery no-clip fish and 2.2-4.7% for total unique fish past Lower Granite Dam.

Weekly estimates of total escapement over Lower Granite Dam tracked the window counts and trap estimates (One example: Figure \@ref(fig:jags-fit-fig). Similar plots for all model runs [each species and year combination] are available on the manuscript GitHub page: [https://www.github.com/KevinSee/ManuscriptSTADEM]([https://www.github.com/KevinSee/ManuscriptSTADEM])). STADEM point estimates were often between estimates based on window counts and those based on the number of fish caught in the adult trap. However, for weeks when very few fish were caught in the trap or there was more uncertainty about the trap rate, STADEM estimates tracked the window counts more closely, as seen in the second week of July 2014, in Figure \@ref(fig:jags-fit-fig). That year also shows the utility of STADEM in dealing with missing data, as the trap was shut down for several weeks in July and August. The model's uncertainty is always smaller than the uncertainty from the trap estimates alone, whereas the window counts alone provide no estimate of uncertainty.

Estimates of weekly nighttime passage and re-ascension rates were not equal in most cases (Figure \@ref(fig:night-reasc-diff-fig)). In particular, there were several weeks when the window counts were quite large, but the nighttime passage and re-ascension rates differed by as much as 10%, in which case assuming them equal would lead to a biased estimate of escapement. When nighttime passage is larger than re-ascension, the window counts will be biased low, and vice versa.

# Discussion

We have presented a novel method for estimating adult salmonid escapement past a large hydroelectric facility (e.g., Lower Granite Dam) that incorporates data from window counts, a fish trap, and observations of PIT tagged fish in the adult passage ladder. Our model explicitly accounts for nighttime passage, re-ascension, and potential error in both window and trap estimates. In doing so, we demonstrated that at Lower Granite Dam, nighttime passage and re-ascension rates do not always offset each other, and assuming they do will lead to biased escapement estimates in some years. With minor adjustments this modeling framework and the STADEM package could be applied to similar migratory species at Lower Granite Dam (e.g., fall-run Chinook Salmon, Pacific lamprey *Lampetra tridentata*), or elsewhere, provided a fish passage barrier with a counting mechanism, a trap that can be used to sample a portion of the run and tag observation or detection infrastructure (e.g., a PIT tag detection array or similar). Our state-space model combined multiple imperfect sources of data to reduce bias in adult escapement estimates and provided quantitative estimates of uncertainty. Accurate population or group abundance estimates and uncertainty accounting for observation and process error can be particularly important when estimates are used for management and conservation decisions such as population viability analyses.

Combining data from the adult fish trap with live and video window counts provides several benefits. First, it allows us to model observer error in the window counts, which is typically unknown. If estimates rely on window counts alone, quantifying observer error is impossible. Capturing and accounting for known sources of error is prudent to minimize management decision risk. Second, by incorporating both sources of information in a state-space framework, STADEM incorporates missing data at either the observation window or adult trap seamlessly. At Lower Granite Dam, the adult trap has been closed for brief or extended periods of time (i.e., days, weeks) intermittently in five of the last ten years, often during peak run times [@USACE2010; @USACE2011; @USACE2012; @USACE2013; @Ogden2016a; @USACE2015; @USACE2016; @USACE2017; @USACE2018; @USACE2019]. Trap closures are typically associated with elevated water temperatures resulting in potential fish handling stress and/or trap malfunctions [@Ogden2016b]. Given predicted Pacific Northwest climate change scenarios [@Zhang2019] trap closures from high water temperatures may become more commonplace in the future, amplifying the need for a modeling framework that accounts for periods of missing data while still capturing estimate uncertainty. Additionally, having a framework in place that accounts for missing periods of data will allow for increased logistic flexibility if, for example, maintenance or construction is needed at the observational window or adult trap.

STADEM could be modified and run on a weekly basis or in near real-time to provide better in-season estimates for fisheries managers. Currently, the only roadblock to this at Lower Granite Dam is the identification of hatchery no-clip fish using genetic tissue samples [@Steele2013; @Steele2019] collected at the adult trap, which currently is completed post-hoc after the trapping season. The inclusion of genetic information typically results in a reduction in wild escapement estimates and an associated increase in hatchery no-clip escapement. However, if in-season management decisions do not require this correction or could accept the potential bias, origin calls at the trap could be used in-season as a first approximation to escapement. Final post-hoc estimates parsed by origin could then be finalized at season’s end. All other data included in this model (e.g., window counts and PIT observations) are otherwise provided in near real-time by DART. Provided the Lower Granite Dam adult fish trap database was updated and available in near real-time, there are minimal obstacles for adapting the STADEM framework to provide in-season estimates of escapement.

Recently, co-managers in the Snake River basin have adopted the STADEM framework to estimate escapement of sp/sum Chinook Salmon and steelhead past Lower Granite Dam, and returning to tributary or population specific spawning areas [@IPTDSW2020; @Kinzer2020]. Estimates of species and origin-specific escapement at Lower Granite Dam, including known uncertainty, are available to further parse into sex- or age-structured escapement estimates [e.g., @Camacho2018; @Schrader2013] that are important for fisheries management and productivity monitoring of wild populations. As an example, STADEM is being applied at Lower Granite Dam to estimate the total unique wild fish migrating past the dam. Estimates of fish passing the dam are then combined with estimated movement or transition probabilities based on PIT tag observations at instream PIT tag detection systems throughout the Snake River basin, similar to @Waterhouse2020, to estimate escapement to Snake River populations and locations throughout the basin [@Orme2019]. Combined, escapement estimates from STADEM and movement probability estimates provide abundance estimates to given tributaries or populations. With sex and age data collected at the adult fish trap [@Hargrove2019], this approach provides necessary information to evaluate productivity and population viability for select Snake River sp/sum Chinook Salmon and steelhead groups [@IPTDSW2020].

Although STADEM was developed with salmonid escapement at Lower Granite Dam in mind, it could be applied to any migratory fish species at locations with similar monitoring infrastructure. Justification and infrastructure exist for applying a modified STADEM framework for fish passing Bonneville Dam, the lowest dam on the Columbia River, or Priest Rapids Dam in the upper Columbia River. Both locations currently trap a sub-sample of passing Chinook Salmon and/or steelhead for biological information and use window counts as a surrogate of true escapement. However, each has at least some similar problems to those observed at Lower Granite Dam such as unaccounted observer and sampling error, nighttime passage, and/or re-ascension. Certainly, estimating an unbiased total return to the entire Columbia River basin (i.e., Bonneville Dam) and Upper Columbia River with uncertainty would benefit managers and decision making.

## Acknowledgements

Funding for this study and development of the STADEM model was partially provided by the Bonneville Power Administration under project 2003-017-00. Special thanks to Darren Ogden and staff at the Lower Granite Dam adult trap for their hard work and diligent data collection. Thank you to personnel at the Idaho Department of Fish and Game, particularly Paul Bunn, Tim Copeland, and Bill Schrader, for providing access to data from the adult fish trap and conceptualizing methods, and to Matthew Campbell and staff at the Eagle Fish Genetics Laboratory for analyzing genetic samples. Special thanks to Columbia Basin Research staff and the Columbia River Data Access in Real Time (DART) application, and Susannah Iltis in particular. And finally, thank you to Rick Orme and Brice Semmens for their contributions to the development of STADEM, as well as Eric Buhle, Sarah Hoffman, John Hargrove and all the other people who have contributed through productive critique and conversations.

\newpage

# References 
<!-- The following line ensures the references appear here for the MS Word or HTML output files, rather than right at the end of the document (this will not work for PDF files):  -->
<div id="refs"></div>

\newpage

# Tables

```{r lgr-sim-tab}
crossing(`Trap rate` = c('0.15', 
                         "0.15 and 0.00 3 weeks"),
         `Fallback & Re-ascension rate` = c(0.06,
                             0.10),
         # `Re-ascension rate` = 1,
         `Window count error` = factor(c("No Error",
                                         "5% Error",
                                         "10% Error"),
                                       levels = c("No Error",
                                                  "5% Error",
                                                  "10% Error"))) %>%
  mutate(`Nighttime passage rate` = if_else(`Fallback & Re-ascension rate` == 0.06,
                                            0.06,
                                            0.05)) %>%
  mutate(Scenario = paste(rep(c('Baseline',
                                "N-R",
                                "N-R trap down",
                                "Trap down"),
                              each = 3)),
         Scenario = if_else(`Window count error` == "5% Error",
                            paste(Scenario, "Err L"),
                            if_else(`Window count error` == "10% Error",
                                    paste(Scenario, "Err H"),
                                    Scenario))) %>%
  select(Scenario, 
         `Trap rate`:`Fallback & Re-ascension rate`,
         `Nighttime passage rate`,
         `Window count error`) %>%
  kable(digits = 2,
        booktabs = T,
        linesep = "",
        caption = "Summary of simulation scenarios including varying adult trapping, fallback and re-ascension, nighttime passage, and window count error rates used to evaluate the performance of STADEM.") %>%
  kable_styling(latex_options = c('scale_down',
                                  "hold_position")) %>%
  column_spec(column = 2,
              width = "1in") %>%
  column_spec(column = c(3, 4, 5),
              width = "1.2in")

```

\newpage

```{r summ-tab}

sum_tab <- res_long_df %>%
  filter(Variable == 'Unique.Wild.Fish',
         Model == 'STADEM') %>%
  group_by(Scenario) %>%
  summarise(inCI = sum(inCI),
            nsims = n(),
            Sample_Mean = mean(est),
            `Relative bias` = mean(rel_bias),
            SE = sqrt(var(est)),
            `Mean CV` = mean(cv),
            RMSE = sqrt(mean(bias^2)),
            Coverage = inCI/nsims) %>%
  mutate(Scenario = str_replace_all(Scenario,
                                    "_", " "),
         Scenario = str_replace(Scenario,
                                "TrapDown",
                                "Trap Down")) %>%
  select(Scenario,
         `Relative bias`,
         `Mean CV`,
         RMSE,
         Coverage)

sum_tab %>%
  kable(digits = c(0, 3, 3, 0, 3),
        booktabs = T,
        linesep = "",
        caption = "Summary statistics, including relative bias, mean coefficient of variation (CV), root mean squared error (RMSE) and 95\\% credible interval coverage for results from each of the twelve simulation scenarios.") %>%
  kable_styling(latex_options = c('scale_down',
                                  "hold_position"))

```

\newpage

```{r lgr-tab}
lgr_est %>%
  mutate(prnt_val = if_else(is.na(cv),
                            as.character(prettyNum(round(mean), big.mark = ',')),
                            paste0(prettyNum(round(mean), big.mark = ","), " (", round(cv, 3), ")"))) %>%
  pivot_wider(id_cols = c("Species",
                          "Year"),
              names_from = "variable",
              values_from = "prnt_val") %>%
  select(Species, 
         Year,
         `Window counts` = win_cnt,
         Total = X.tot.new.all,
         Wild = X.tot.new.wild,
         Hatchery = X.tot.new.hatch,
         `Hatchery no-clip` = X.tot.new.hnc) %>%
  kable(booktabs = T,
        linesep = "",
        caption = "Window counts, and estimates of total, wild, hatchery and hatchery no-clip escapement, with coefficients of variation in parenthesis, for spring/summer-run Chinook Salmon and steelhead from spawn years 2010 to 2019.") %>%
  kable_styling(latex_options = c('scale_down',
                                  "hold_position"))

```

\newpage

\listoffigures

\newpage

# Figures

```{r stadem-examp-fig, fig.height = 5, fig.cap = "Schematic showing the STADEM model framework. (A) shows the posterior of the estimate of fish crossing the dam while the window is open (dashed line shows observed window counts). That estimate is divided by the nighttime passage rate (B). The total fish is then discounted by the re-ascension rate to estimate unique fish (C). Those unique fish are then multiplied by the proportions by origin (D), to estimate unique fish by origin."}

arrow_length = unit(0.15, 'inches')
arrow_height = 1.05
my_xlim = c(-4, 7.5)
my_ylim = c(0.08, 1.2)

# choose appropriate color scale for color vs. black & white
p1 = ggplot(data.frame(x = c(-5, 8)),
            aes(x)) +
  stat_function(fun = dnorm,
                args = list(mean = 4, sd = 0.4),
                aes(color = 'Daytime Fish'),
                lwd = 2) +
  stat_function(fun = dnorm,
                args = list(mean = 6, sd = 0.55),
                aes(color = 'Total Fish'),
                geom = 'blank') +
  stat_function(fun = dnorm,
                args = list(mean = 5, sd = 0.6),
                aes(color = 'Unique Fish'),
                geom = 'blank') +
  stat_function(fun = dnorm,
                args = list(mean = -1, sd = 0.75),
                aes(color = 'Unique Wild Fish'),
                geom = 'blank') +
  stat_function(fun = dnorm,
                args = list(mean = -1, sd = 0.75),
                aes(color = 'Unique Hatchery Fish'),
                geom = 'blank') +
  stat_function(fun = dnorm,
                args = list(mean = -1, sd = 0.75),
                aes(color = 'Unique HNC Fish'),
                geom = 'blank') +
  scale_color_brewer(palette = "Spectral") +
  geom_vline(xintercept = 4,
             linetype = 2) +
  theme_pubr(legend = "bottom") +
  guides(color = guide_legend(nrow = 3,
                              title.position = 'top')) +
  theme(axis.text = element_blank(),
        axis.ticks = element_blank(),
        axis.title.y = element_blank()) +
  labs(x = 'Escapement',
       color = 'Fish Type',
       title = '   Daytime Fish') +
  coord_cartesian(xlim = my_xlim,
                  ylim = my_ylim)
# coord_fixed(ratio = 4,
  #             xlim = my_xlim,
  #             ylim = my_ylim)

p2 = p1 +
  stat_function(fun = dnorm,
                args = list(mean = 6, sd = 0.55),
                aes(color = 'Total Fish'),
                lwd = 2) +
  labs(title = '   Nighttime Passage')

p2a = p2 +
  geom_segment(x = 4,
               y = arrow_height,
               xend = 6,
               yend = arrow_height,
               arrow = arrow(length = arrow_length))

p3 = p2 +
  stat_function(fun = dnorm,
                args = list(mean = 5, sd = 0.6),
                aes(color = 'Unique Fish'),
                lwd = 2) +
  labs(title = '   Re-ascension')

p3a = p3 +
  geom_segment(x = 6,
               y = arrow_height,
               xend = 5,
               yend = arrow_height,
               arrow = arrow(length = arrow_length))

p4 = p3 +
  stat_function(fun = dnorm,
                args = list(mean = -1, sd = 0.85),
                aes(color = 'Unique Wild Fish'),
                lwd = 2) +
  stat_function(fun = dnorm,
                args = list(mean = 1, sd = 0.75),
                aes(color = 'Unique Hatchery Fish'),
                lwd = 2) +
  stat_function(fun = dnorm,
                args = list(mean = -2, sd = 1.25),
                aes(color = 'Unique HNC Fish'),
                lwd = 2) +
  geom_segment(x = 5,
               y = arrow_height,
               xend = -1,
               yend = arrow_height,
               arrow = arrow(length = arrow_length)) +
  labs(title = '   Origin Proportion')

examp_col = ggarrange(p1, p2a, p3a, p4,
                    labels = 'AUTO',
                    ncol = 2,
                    nrow = 2,
                    align = 'v',
                    common.legend = T,
                    legend = 'bottom')

bw_scale = c(0.9, 0.1)
examp_bw = ggarrange(p1 +
                       scale_color_grey(start = bw_scale[1],
                                        end = bw_scale[2]), 
                   p2a +
                       scale_color_grey(start = bw_scale[1],
                                        end = bw_scale[2]), 
                   p3a +
                       scale_color_grey(start = bw_scale[1],
                                        end = bw_scale[2]), 
                   p4 +
                       scale_color_grey(start = bw_scale[1],
                                        end = bw_scale[2]),
                    labels = 'AUTO',
                    ncol = 2,
                    nrow = 2,
                    align = 'v',
                    common.legend = T,
                    legend = 'bottom')

examp_bw

# save black and white version
ggsave(paste0("../figures/stadem-examp-fig-bw.", my_ext),
       examp_bw,
       height = 5,
       width = 7,
       dpi = my_dpi)

# save color version
ggsave(paste0("../figures/stadem-examp-fig-col.", my_ext),
       examp_col,
       height = 5,
       width = 7,
       dpi = my_dpi)

```

\newpage

```{r rel-bias-lgd, fig.cap = "Boxplots of the relative bias of STADEM estimates for wild escapement across various scenarios (See Table 1)."}
# hdi_box = function(x) {
#   r = c(hdi(x, 0.95), hdi(x, 0.5), median(x))
#   # r = quantile(x, probs = c(0.025, 0.975, 0.25, 0.75, 0.5))
#   names(r) = c('ymin', 'ymax', 'lower', 'upper', 'middle')
#   r
# }
# 
# hdi_out = function(x) {
#   subset(x, x < hdi(x, 0.95)[1] | x > hdi(x, 0.95)[2])
#   # subset(x, x < quantile(x, probs = c(0.025)) | x > quantile(x, probs = c(0.975)))
# }

rel_bias_p = res_long_df %>%
  filter(grepl('Unique', Variable),
         Model == 'STADEM') %>%
  mutate(Variable = fct_recode(Variable,
                               Wild = 'Unique.Wild.Fish',
                               Hatchery = 'Unique.Hatch.Fish',
                               HNC = 'Unique.HNC.Fish')) %>%
  mutate(grp = paste(Variable, Model, sep = ' / ')) %>%
  filter(Variable == "Wild") %>%
  ggplot(aes(x = rel_bias,
             y = fct_rev(Scenario),
             fill = grp)) +
  # stat_summary(fun.data = hdi_box,
  #              geom = 'boxplot') +
  # stat_summary(fun = hdi_out,
  #              geom = 'point',
  #              color = 'gray') +
  geom_boxplot(fill = 'gray80') +
  geom_vline(xintercept = 0,
             linetype = 2) +
  # facet_wrap(~ Variable) +
  theme_pubr(legend = "none") +
  theme(axis.text = element_text(size = 10)) +
  labs(x = 'Relative Bias',
       y = "Scenario",
       fill = 'Fish / Model') +
  coord_fixed(0.03)

rel_bias_p

# save black and white version
ggsave(paste0("../figures/rel-bias-lgd-bw.", my_ext),
       rel_bias_p,
       height = 5,
       width = 3.5,
       dpi = my_dpi)

```

\newpage

```{r jags-fit-fig, fig.cap = "Time-series plot showing estimates of total escapement for spring/summer-run Chinook Salmon in 2014, including window counts, trap estimates and STADEM estimates. The dark gray ribbon represents the 95% credible intervals for STADEM estimates, while the light gray ribbon represents the 95% confidence intervals for the trap estimates."}

# get plotting data from particular year
plot_df = tibble(Species = "Chinook",
                 Year = "2014") %>%
  mutate(plot_title = paste(Species, Year)) %>%
  rowwise() %>%
  mutate(plot_data = map2(Species, 
                          Year,
                          .f = function(x, y) {
                            load(paste0('../data/derived_data/STADEM_results/LGR_STADEM_', Species, '_', Year, '.rda'))
                            plot_df = stadem_mod$summary[grep('^X.all', rownames(stadem_mod$summary)),] %>%
                              as_tibble(rownames = 'var') %>%
                              mutate(week = as.integer(str_extract(var, "[0-9]+")),
                                     param = str_extract_all(var, "[:alpha:]+", simplify = T)[,3],
                                     param = ifelse(param == '', 'all', param)) %>%
                              select(var, param, week, everything()) %>%
                              left_join(stadem_list$weeklyData %>%
                                          rename(week = week_num) %>%
                                          rowwise() %>%
                                          mutate(trap_lowCI = max(0, trap_est + trap_est_se * qnorm(0.025)),
                                                 trap_uppCI = trap_est + trap_est_se * qnorm(0.975)) %>%
                                          ungroup())%>%
                              mutate_at(vars(trap_est),
                                        list(~ if_else(. == Inf,
                                                       as.numeric(NA),
                                                       .))) %>%
                              select(-Species)
                            return(plot_df)
                          })) %>%
  unnest(cols = plot_data) %>%
  select(Species, Year, plot_title, everything()) %>%
  filter(param == "all")

# drop initial weeks when window counts == 0
plot_df %<>%
  group_by(Species, Year) %>%
  filter(Start_Date >= min(Start_Date[win_cnt > 0]))

# scale some variables for plotting
plot_df %<>%
    mutate_at(vars(trap_lowCI,
                 trap_uppCI,
                 `2.5%`,
                 `97.5%`,
                 win_cnt,
                 trap_est,
                 `50%`),
            list(~ . / 10000))

# set size of points
pt_size = 4

# color version
ts_col = plot_df %>%
  ggplot(aes(x = Start_Date)) +
  geom_ribbon(aes(ymin = trap_lowCI,
                  ymax = trap_uppCI,
                  fill = "Trap Est."),
              alpha = 0.4) +
  geom_ribbon(aes(ymin = `2.5%`,
                  ymax = `97.5%`,
                  fill = "Model"),
              alpha = 0.4) +
  geom_point(aes(y = win_cnt,
                 shape = 'Window Count',
                 color = 'Window Count'),
             size = pt_size) +
  geom_point(aes(y = trap_est,
                 shape = 'Trap Est.',
                 color = 'Trap Est.'),
             size = pt_size) +
  geom_line(aes(y = `50%`,
                color = 'Model')) +
  geom_point(aes(y = `50%`,
                 shape = 'Model',
                 color = 'Model'),
             size = pt_size) +
  scale_color_manual(values = c('Model' = 'black',
                                'Window Count' = 'blue',
                                'Trap Est.' = 'red'),
                     name = "Source") +
  scale_fill_manual(values = c('Model' = 'black',
                                'Window Count' = 'blue',
                                'Trap Est.' = 'red'),
                     guide = "none") +
  scale_shape_manual(values = c("Model" = 18,
                                "Window Count" = 1,
                                "Trap Est." = 8),
                     name = "Source") +
  theme_pubr(legend = "bottom") +
  theme(panel.background = element_rect(color = "transparent"),
        plot.background = element_rect(color = "transparent"),
        legend.background = element_rect(color = "transparent")) +
  facet_wrap(~ plot_title,
             scales = "free",
             ncol = 2) +
  labs(x = 'Date',
       y = 'Total Escapement (x 10,000)')

# black and white version
ts_bw = ts_col +
  scale_color_grey(start = 0,
                   end = 0,
                   name = "Source") +
  scale_fill_manual(values = c('Model' = 'gray20',
                                'Trap Est.' = 'gray70'),
                     guide = "none")

# ts_col
ts_bw

# save black and white version
ggsave(paste0("../figures/jags-fit-fig-bw.", my_ext),
       ts_bw,
       height = 5,
       width = 7,
       dpi = my_dpi)

# save color version
ggsave(paste0("../figures/jags-fit-fig-col.", my_ext),
       ts_col,
       height = 5,
       width = 7,
       dpi = my_dpi)
```

\newpage

```{r night-reasc-diff-fig, fig.height=5, fig.cap = "Nighttime passage rate plotted against re-ascension rate on the logit scale, calculated from observed PIT tags for each week of spawn years 2010-2019. The size of each point is proportional to the window count that week. The dashed line is the 1-1 line."}

rate_p_base = rate_comp %>%
  filter(win_cnt > 0) %>%
  filter(tot_tags > 0,
         !night_rate %in% c(1, 0),
         !reasc_rate %in% c(1, 0)) %>%
  ggplot(aes(x = night_rate,
             y = reasc_rate)) +
  geom_abline(linetype = 2) +
  facet_wrap(~ Species,
             ncol = 1) +
  theme_pubr(legend = 'none') +
  coord_fixed(ratio = 1) +
  scale_x_continuous(trans = "logit",
                     breaks = scales::pretty_breaks(3)) +
  scale_y_continuous(trans = "logit",
                     breaks = scales::pretty_breaks()) +
  labs(x = "Nighttime Passage Rate",
       y = "Re-ascension Rate")

# color version
rate_p_col = rate_p_base +
  scale_color_brewer(palette = "Paired") +
  geom_point(aes(size = win_cnt / 1000,
                 color = as.factor(Year))) +
  labs(color = "Spawn Year",
       size = "Window Count\n (x 1,000)") +
  theme_pubr(legend = "left") +
  theme(axis.text = element_text(size = 10))

# black and white version
rate_p_bw = rate_p_base +
  geom_point(aes(size = win_cnt / 1000),
             shape = 21,
             fill = 'gray40',
             color = 'black',
             alpha = 0.3) +
  labs(size = "Window Count\n (x 1,000)") +
  theme_pubr(legend = "bottom") +
  theme(axis.text = element_text(size = 10)) +
  guides(size = guide_legend(nrow = 2))

# rate_p_col
rate_p_bw

# save black and white version
ggsave(paste0("../figures/night-reasc-diff-fig-bw.", my_ext),
       rate_p_bw,
       width = 3.5,
       height = 7,
       dpi = my_dpi)

# save color version
ggsave(paste0("../figures/night-reasc-diff-fig-col.", my_ext),
       rate_p_col,
       width = 5,
       height = 7,
       dpi = my_dpi)
```

\newpage

```{r child = 'appendices/Appendix_STADEM_descrip.Rmd'}
```

\newpage

```{r child = 'appendices/Appendix_simulation_details.Rmd'}
```